DATA 621 - Business Analytics and Data Mining

### 2020-11-26 HW5/HW5-Draft.Rmd
Minor copyediting and specifying amsmath LaTeX package

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #5'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/28/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 4
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

@@

The right balance of the chemical properties of the wine is essential to make it
taste, look, and smell delightful, in addition to having a long shelf-life.
Typically, citric acid and alcohol are the main positive attributes a wine
should have, while volatile acidity and sulphates are the main negative ones. 
However, the dataset has negative values instead of near-zero in some of these 
variables describing the chemical properties of the wine. Because little
information is given whether or not these variables were normalized, absolute
values will be taken assuming there was an error made during data collection.

@@

```{r m1PNB}
m1PNBC <- data.table(Model = c("Poisson 1", "Negative Binomial 1"),
                     Deviance = c(m1PFit$finalModel$deviance,
                                  m1NBFit$deviance),
                     AIC = c(m1PFit$finalModel$aic,
                             m1NBFit$aic))
kable(m1PNBC, digits = 3L,
      caption = "Poisson 1 and NegBinom 1 Goodness-of-Fit Comparison")
```

## Poisson Model #2 & Multiple Linear Regrssion Model #1
The `TARGET` variable has 2734 observations with zero values. This could be a
result of the manufacturer not releasing the wine into the market, or that there
were no orders placed. Therefore, backward stepwise variable elimination for a
zero-inflated poisson (ZIP) analysis will be investigated to understand the
conditions. This model is further compared to the same Poisson model to
determine whether there is an improvement.

@@

Therefore, the zero-inflated poisson model have two parameters, \(\lambda_i\)
and \(\omega_i\), and linked with the predictor variables with the following:

\[
\begin{align}
&\log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji}\\
&\log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji}
\end{align}
\]

@@

\[
\begin{align}
\mathrm{E}(Y_i) &= \lambda_i(1 - \omega_i)\\
\mathrm{Var}(Y_i) &= (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2)
\end{align}
\]

@@

\small
```{r m2PTable}
kable(summary(m2.fpmodel)$coefficients$count,
      caption = "Zero Inflated Poisson Model #2 Count Output")
kable(summary(m2.fpmodel)$coefficients$zero,
      caption = "Zero Inflated Poisson Model #2 Zero Output")
```
\normalsize

@@

The zero model contains information about variables that the nested Logistic 
Regression model has been used to estimate the probability of whether or not cases are
purchased. Examining the variables with a negative regression coefficient, this
suggests that as the factor for the specific chemical properties increases, the
probability of no case being purchase decreases. This is in line with our intuition,
particularly for the sulfur dioxide levels. When a wine has an unusually small 
percentage of the Free and Total $SO_2$, this suggests the wine is chemically and/or 
microbially unstable. Thus, if a wine manufacturer were to increase the sulfur 
dioxide levels by one unit, the odds that the wine would certainly not be purchased 
would decrease by a factor of 0.92 to 0.95. However, `LabelAppeal` is among the
positive coefficients for the zero model. This is interpreted as the higher a 
rating is for a bottle of wine, the more likely there will be no case purchase, and it
doesn't seem reasonable.

### Multiple Linear Regrssion Model #1

Multiple Linear Regression Model #1 will be using the step function has an options 
to use a process that both adds and removes variables in the model. It uses AIC 
(Akaike information criterion) as a selection criterion. A full model and a null 
are defined, and then the function will follow a procedure to find the model with 
the lowest AIC.

@@

\small
```{r m1MLRTable}
kable(summary(m1MLRFit)$coefficient,
      caption = "Multiple Linear Regression Model #1 Output")
```
\normalsize

@@

## Negative Binomial Model #2 & Multiple Linear Regrssion Model #2

### Negative Binomial Model #2
#### Baseline Model 
We will start with a simple negative binomial regression model to serve as a 
baseline. This includes all variables in the dataset.

```{r m2NBTrainBase}
m2NBBase <- glm.nb(TARGET ~ ., data = m3train)
knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Base Output')
```
Immediately, we can see a few variables that do not meet the 0.05
p-value threshold.

@@

```{r m2NBTrain}
m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
                LabelAppeal + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Final Output')
```

@@

### Multiple Linear Regrssion Model # 2
One drawback to traditional multiple linear regression models is that the 
forecasted counts are not integers. We will deal with this by rounding the
predictions to the nearest whole number.

@@

```{r m2MLTrainBase}
m2MLBase <- lm(TARGET ~ ., data = m3train)
knitr::kable(summary(m2MLBase)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Base Output')
```
Most of the variables in the dataset are significant, but we can immediately
see a few that are above the 0.05 p-value threshold.

@@

```{r m2MLTrain}
m2MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide 
              + TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal 
              + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2MLFit)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Final Output')
```

>>>>

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #5'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/28/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 4
    extra_dependencies:
      amsmath: null
      inputenc: utf8
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

@@

The right balance of the chemical properties of the wine is essential to make it
taste, look, and smell delightful, in addition to having a long shelf-life.
Typically, citric acid and alcohol are the main positive attributes a wine
should have, while volatile acidity and sulfates are the main negative ones. 
However, the dataset has negative values instead of near-zero in some of these 
variables describing the chemical properties of the wine. Because little
information is given whether or not these variables were normalized, absolute
values will be taken assuming there was an error made during data collection.

@@

```{r m1PNB}
# P1 & NB1: GoF Comparison
m1PNBC <- data.table(Model = c("Poisson 1", "Negative Binomial 1"),
                     Deviance = c(m1PFit$finalModel$deviance,
                                  m1NBFit$deviance),
                     AIC = c(m1PFit$finalModel$aic,
                             m1NBFit$aic))
kable(m1PNBC, digits = 3L,
      caption = "Poisson 1 and NegBinom 1 Goodness-of-Fit Comparison")
```

## Poisson Model #2 & Multiple Linear Regression Model #1
The `TARGET` variable has 2734 observations with zero values. This could be a
result of the manufacturer not releasing the wine into the market, or that there
were no orders placed. Therefore, backward stepwise variable elimination for a
zero-inflated Poisson (ZIP) analysis will be investigated to understand the
conditions. This model is further compared to the same Poisson model to
determine whether there is an improvement.

@@

Therefore, the zero-inflated Poisson model have two parameters, \(\lambda_i\)
and \(\omega_i\), and linked with the predictor variables with the following:

\[
\begin{aligned}
&\log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji}\\
&\log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji}
\end{aligned}
\]

@@

\[
\begin{aligned}
\mathrm{E}(Y_i) &= \lambda_i(1 - \omega_i)\\
\mathrm{Var}(Y_i) &= (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2)
\end{aligned}
\]

@@

\small
```{r m2PTable}
# P2: Outputs
kable(summary(m2.fpmodel)$coefficients$count,
      caption = "Zero Inflated Poisson Model #2 Count Output")
kable(summary(m2.fpmodel)$coefficients$zero,
      caption = "Zero Inflated Poisson Model #2 Zero Output")
```
\normalsize

@@

The zero model contains information about variables that the nested Logistic 
Regression model has been used to estimate the probability of whether or not
cases are purchased. Examining the variables with a negative regression
coefficient, this suggests that as the factor for the specific chemical
properties increases, the probability of no case being purchase decreases. This
is in line with our intuition, particularly for the sulfur dioxide levels. When
a wine has an unusually small percentage of the Free and Total \(SO_2\), this
suggests the wine is chemically and/or microbially unstable. Thus, if a wine
manufacturer were to increase the sulfur dioxide levels by one unit, the odds
that the wine would certainly not be purchased would decrease by a factor of
0.92 to 0.95. However, `LabelAppeal` is among the positive coefficients for the
zero model. This is interpreted as the higher a rating is for a bottle of wine,
the more likely there will be no case purchase, and it doesn't seem reasonable.

### Multiple Linear Regrssion Model #1
Multiple Linear Regression Model #1 will be using the step function has an
option to use a process that both adds and removes variables in the model. It
uses AIC (Akaike information criterion) as a selection criterion. A full model
and a null are defined, and then the function will follow a procedure to find
the model with the lowest AIC.

@@

\small
```{r m1MLRTable}
# ML1: Outputs
kable(summary(m1MLRFit)$coefficient,
      caption = "Multiple Linear Regression Model #1 Output")
```
\normalsize

@@

## Negative Binomial Model #2 & Multiple Linear Regrssion Model #2
### Negative Binomial Model #2
#### Baseline Model 
We will start with a simple negative binomial regression model to serve as a 
baseline. This includes all variables in the dataset.

```{r m2NBTrainBase}
# NB2: Baseline Model
m2NBBase <- glm.nb(TARGET ~ ., data = m3train)
knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Base Output')
```

Immediately, we can see a few variables that do not meet the 0.05 p-value
threshold.

@@

```{r m2NBTrain}
# NB2: Revised Model
m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
                LabelAppeal + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Final Output')
```

@@

### Multiple Linear Regrssion Model # 2
One drawback to traditional multiple linear regression models is that the 
forecast counts are not integers. We will deal with this by rounding the
predictions to the nearest whole number.

@@

```{r m2MLTrainBase}
# ML2: Baseline model
m2MLBase <- lm(TARGET ~ ., data = m3train)
knitr::kable(summary(m2MLBase)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Base Output')
```

Most of the variables in the dataset are significant, but we can immediately see
a few that are above the 0.05 p-value threshold.

@@

```{r m2MLTrain}
# ML2: Enhanced model
m2MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide 
              + TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal 
              + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2MLFit)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Final Output')
```
